Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-10 paper/main.tex
Updates

\begin{figure}[t!]
	\centering
	\includegraphics[page=1]{fig5.pdf}
	\caption{DAG learned from the adult income dataset using an LLM (GEMINI-1.5-flash) as the expert. The p-value threshold used is $ 0.05 $ and the measure of association threshold is $ 0. 1 $}
	\label{fig:adult_llm}
\end{figure}
Recently, there has been a lot of work towards using LLMs for causal discovery,
with tasks ranging from determining pairwise edge orientation
\citep{Kiciman2023, Jin2024} to full causal structure learning \citep{Naik2023,
Vashishtha2023} and counterfactual reasoning tasks\citep{Kiciman2023} (see
\citet{Liu2024} for an overview).

Since our approach utilizes expert knowledge to determine edge orientations, we
investigated the potential of using an LLM for this task. We applied our causal
discovery procedure with using Gemini 1.5 flash model as the expert. Our method
uses a greedy approach where the pair of variables with the highest unexplained
correlation is oriented first. The prompt used to determine the orientation is
provided in Appendix~\ref{section:llms}. Our LLM prompt utilizes variable
description to determine the edge orientation. We applied this approach to the
adult income dataset and the output is shown in Fig.~\ref{fig:adult_llm}.

Unlike the other approaches to utilize LLM for doing full causal discovery, we 
used a combination of statistical tests and LLMs. This greatly reduces the
amount of information that we need from the LLM. Our approach is also able to 
ask exact questions about the variables unlike some of the other approaches.

>>>>

\begin{figure}[t!]
	\centering
	\includegraphics[page=1]{fig5.pdf}
	\caption{DAG learned from the adult income dataset using an LLM (GEMINI-1.5-flash) as the expert. The p-value threshold used is $ 0.05 $ and the measure of association threshold is $ 0. 1 $}
	\label{fig:adult_llm}
\end{figure}

Recently, there has been significant interest in leveraging Large Language
Models (LLMs) for causal discovery. These applications range from determining
pairwise edge orientations \citep{Kiciman2023, Jin2024} to full causal
structure learning \citep{Naik2023, Vashishtha2023} and counterfactual
reasoning \citep{Kiciman2023} (see \citet{Liu2024} for a comprehensive
overview).

Since our approach relies on expert knowledge to determine edge orientations,
we explored the potential of using an LLM for this task. We applied our causal
discovery procedure using the Gemini 1.5 Flash model as the expert. We simulate
the \emph{weak ancestral oracle} using the LLM by asking it to choose the
causal direction between a pair of variables. The LLM is provided a description
of each of the variables for the task. We have provided the prompt used in
Appendix~\ref{section:llms}. Figure~\ref{fig:adult_llm} shows the DAG learned
on the Adult Income dataset using the LLM as the expert. Compared to other
methods that use LLMs for causal discovery, our approach combines a data-driven
method with LLMs. We believe this should help in reducing the amount of
mistakes made by the LLM as well as reduce the amount of information needed
from the LLM.
